\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2010-11-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
minor restructuring of examples files; adding flag USE_SOURCES_RECVS_Z in constants.h for locating source/receivers in randomly shaped meshes; updates some files to avoid compiler warnings and unifies character lengths for LOCAL_PATH; updates manual

Since then it has been developed and maintained by a development team: in alphabetical order,
Emanuele Casarotti,
Min Chen,
Vala Hj\"orleifsd\'ottir,
Emiljana Jorgji,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Roland Martin,
Dennis McRitchie,
David Mich\'ea,
Daniel Peter,
Brian Savage,
Leif Strand,
Carl Tape,
Jeroen Tromp.\\

@@

In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
See \texttt{examples/noise\_tomography/README\_NOISE\_TOMOGRAPHY} for explanations. \\

@@

\texttt{./pre-processing}\\

Even if this script does not work on your workstation, the procedure it describes is universal.
You may review the whole process described in the last section by following the commands in \texttt{pre-processing},
which should contain enough explanations for all the commands.
%>YANGL

@@

>>>>

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
Ebru Bozdag,
Emanuele Casarotti,
Min Chen,
Hom Nath Gharti,
Vala Hj\"orleifsd\'ottir,
Emiljana Jorgji,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Dennis McRitchie,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Brian Savage,
Leif Strand,
Carl Tape,
Jeroen Tromp.\\

@@

In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
See \texttt{examples/noise\_tomography/README} for explanations. \\

@@

\texttt{./pre-processing.sh}\\

Even if this script does not work on your workstation, the procedure it describes is universal.
You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
which should contain enough explanations for all the commands.
%>YANGL

@@

The SPECFEM3D software code internally uses Cartesian coordinates.
%This requires to convert the given locations for sources and receiver locations.
Sources and receiver locations are read in from the
\texttt{CMTSOLUTION} and \texttt{STATIONS} file.
Note that e.g. the \texttt{CMTSOLUTION} file denotes the location by "longitude/latitude/depth". We read in
longitude as $x$ coordinate, latitude as $y$ coordinate.
In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
using a UTM conversion with the specified UTM zone.
The value for depth (given in $km$ in \texttt{CMTSOLUTION}) or burial depth (given in $m$ in \texttt{STATIONS})
is evaluated with respect to the surface of the mesh at the specified $x$/$y$ location to find the corresponding $z$ coordinate.
It is possible to use this depth value directly as $z$ coordinate by
changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.

The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
For the labeling of the seismograms, we name "$x$/$y$/$z$" to "BHE/BHN/BHZ" components.
Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
\texttt{.true.}) as vertical. In this case, the outputs are rotated to have the BHZ component normal to the surface of the mesh, BHE and BHN are somewhat arbitrary orthogonal directions along the surface.

\chapter{\label{cha:Troubleshooting}Troubleshooting}

\section*{FAQ}

\begin{description}
\item [configuration fails:]

   Examine the log file 'config.log'. It contains detailed informations.
   In many cases, the path's to these specific compiler commands F90,
   CC and MPIF90 won't be correct if `./configure` fails.

   Please make sure that you have a working installation of a Fortran compiler,
   a C compiler and an MPI implementation. You should be able to compile this
   little program code:

      program main ~\\
        include 'mpif.h' ~\\
        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
        integer ier ~\\
        call MPI\_INIT(ier) ~\\
        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
        call MPI\_FINALIZE(ier) ~\\
      end

\item [compilation fails stating:] ~\\
\begin{lyxcode}
    ...  ~\\
    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
    ...  ~\\
\end{lyxcode}

  Make sure you're pointing to the right 'mpif90' wrapper command.

  Normally, this message will appear when you're mixing two different Fortran
  compilers. That is, using e.g. gfortran to compile non-MPI files
  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.

  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
